---
title: "testsheets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testsheets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# testsheets

## Overview

`testsheets` is a package for creating [testthat](https://testthat.r-lib.org/index.html) files from spreadsheets. It supports Google Sheets and Excel documents and relies heavily on the [tidyverse](https://www.tidyverse.org/) packages [googlesheets4](https://googlesheets4.tidyverse.org) and [readxl](https://readxl.tidyverse.org/index.html).

## Installation

This package is can be installed from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nicole-brewer/testsheets")
```

## Terminology

Spreadsheet terminolgies among major tech giants are confusing and naming conflicts is worth extricating. In this package, we use three terms that conform to naming convensions in the **googlesheets4** package.
- **sheet**:  a singular, two dimensional set of tabular data, also called a "sheet" in Google Sheets, and "spreadsheet" and "worksheet" in Excel
- **ss**: a document that contains one or more sheets, called a "spreadsheet" in GoogleSheets, and "workbook" in Excel
- **testsheet**: a **sheet** of any mimetype that follows formatting rules that allow for conversion a **testfile**
- **testfile**: a script created from a **testsheet** that is compatable with the **testthat** package

## Basic Usage

Let's say we want to test a function we have written. In this example we will use a Google Sheets document called "example" that has two sheets named after the functions they were designed to test: "sum" and "mean".

```{r}
# Retrieve the google sheet
library(googledrive)
library(testsheets)

# indicate there is no need for an access token since we are accessing
# a read-only sheet via a public link
googledrive::drive_deauth()

# get 
sum_ss <- googledrive::drive_get("https://docs.google.com/spreadsheets/d/1vR1aUlKMGtZIJOZaHNk7kePgeYYrkhoy0PFmZxs8wt4/edit?usp=sharing")

# Load the "sum" sheet into a dribble
create_testfile(sum_ss, sheet = "sum")
```

## Advanced Usage: Leveraging testdata

Internally, `create_testfile` obtains a tibble containing `testsheet` cell data. This tibble is cleaned in certain ways: for example, rows that have "false" in the include column will be removed and the 'param_' prefixes indicating input parameters will dropped.

Perhaps you have a lot of injected code in your spreadsheet that needs cleaning or repair to create an error_free **testfile**. `create_testfile` is not an ideal funciton in this scenario, because you skip over the intermediary tibble that we will call **testdata**.

`testsheets` provides a workaround by providing two functions that separate testfile creation into two steps: `read_testsheet` and `write_testfile`. 

```{r}
library(testsheets)

# Obtain the path of an example testsheet
path <- testsheets_example("excel")

# create testdata from the sheet called "sum"
testdata <- read_testsheet(path, sheet = "sum")

# modify the testdata tibble in some desired way
change_testdata <- function(testdata) {
  message("Testdata has been modified in some way!")
  testdata <- testdata
}

testdata <- testdata %>% change_testdata()

# write the testdata to a testfile
write_testfile(testdata, "sum")
```
### Cleaning up the advanced workflow

Now that we have written and understood whats going on behind the scenes in an advanced workflow, we have the option of cleaning it up by reverting back to the `create_testfile()` function used in the basic workflow, but with one extra parameter. In R, functions are objects (of class type "function") just like anything else, so we can pass our custom `change_testdata` function as a `manipulator` parameter.

```{r}
library(testsheets)

# Obtain the path of an example testsheet
path <- testsheets_example("excel")

# define our "manipulator" function
change_testdata <- function(testdata) {
  message("Testdata has been modified in some way!")
  testdata <- testdata
}

create_testfile(path, "sum", manipulator=change_testdata)
```




