---
title: "testsheets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testsheets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# testsheets

## Overview

`testsheets` is a package for creating [testthat](https://testthat.r-lib.org/index.html) files from spreadsheets. It supports Google Sheets and Excel documents and relies heavily on the [tidyverse](https://www.tidyverse.org/) packages [googlesheets4](https://googlesheets4.tidyverse.org) and [readxl](https://readxl.tidyverse.org/index.html).

## Installation

This package is can be installed from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("nicole-brewer/testsheets")
```

## Terminology

Spreadsheet terminolgies among major tech giants are confusing and naming conflicts is worth extricating. In this package, we use three terms that conform to naming convensions in the **googlesheets4** package.
- **sheet**:  a singular, two dimensional set of tabular data, also called a "sheet" in Google Sheets, and "spreadsheet" and "worksheet" in Excel
- **ss**: a document that contains one or more sheets, called a "spreadsheet" in GoogleSheets, and "workbook" in Excel
- **testsheet**: a **sheet** of any mimetype that follows formatting rules that allow for conversion a **testfile**
- **testfile**: a script created from a **testsheet** that is compatable with the **testthat** package


## Example

Let's say we want to test a function we have written. In this example we will use a Google Sheets document called "example" that has two sheets named after the functions they were designed to test: "sum" and "mean".

```{r}
# Retrieve the google sheet
library(googledrive)
library(testsheets)

# indicate there is no need for an access token since we are accessing
# a read-only sheet via a public link
googledrive::drive_deauth()

# get 
sum_ss <- googledrive::drive_get("https://docs.google.com/spreadsheets/d/1vR1aUlKMGtZIJOZaHNk7kePgeYYrkhoy0PFmZxs8wt4/edit?usp=sharing")

# Load the "sum" sheet into a dribble
#create_testsheet(sum_ss, sheet = "sum")
```



## Advanced Workflow: Leveraging testdata

Internally, `create_testfile` obtains a tibble containing `testsheet` cell data. This tibble is cleaned in certain ways: for example, rows that have "false" in the include column will be removed and the 'param_' prefixes indicating input parameters will dropped.

Perhaps you have a lot of injected code in your spreadsheet that needs cleaning or repair to create an error_free **testfile**. `create_testsheet` is not an ideal funciton in this scenario, because you skip over the intermediary tibble that we will call **testdata**.

`testsheets` provides a workaround by providing two functions that separate testfile creation into two steps: `read_testsheet` and `write_testfile`. 

```{r}


# create testdata from temporary xlsx file
#testdata <- read_testsheet(sum_ss, sheet = "sum")

# modify the testdata tibble in some desired way
change_testdata <- function(testdata) {
  print('make modifications here!')
  testdata <- testdata
}
#testdata <- testdata %>% change_testdata()

# write the testdata to a testfile
#write_testsheet(testdata)
```
* Side note: `readxl` doesn't currently support url's as file locations. See [issue 278](https://github.com/tidyverse/readxl/issues/278).





